The Talking Heads Experiment

### 2015-04-16 chap3/chap3.tex
hyphenation, orphans, widows and some other layout updated

So we find that the world does not present itself neatly segmented, 
processed, and categorised. There is an enormous 
gap between the symbolic world of objects and categories
and the subsymbolic world of analog sensing. 
The brain somehow performs a vast amount of processing
to fill this gap, without us being in the least aware of it. 
This processing takes place in parallel. Many different
segmentation procedures and feature detectors 
operate concurrently on streams of consecutive images, 
generating hypotheses that become stronger if they are
confirmed by additional evidence, or weaker if 
they do not fit into a larger picture. We are only 
aware of the final result and therefore have no 
intuition of what is really going on, except
in rare and pathological circumstances. 
Rather than viewing perception as a straight-forward step-by-step
transformation of the raw sensory data into a segmented
picture, it is better to think of the whole process
as a boiling soup with thousands of hypotheses taking
shape, some of them floating like bubbles up to the surface. Constraints
and expectations flow down from the top so that the maximum 
amount of available information is used to construct the 
coherent segmented picture of reality we consciously experience.

@@

Sensors and actuators are the interfaces between the physical
world and the internal world.\is{sensors}\is{actuators} 
They are dedicated hard-wired components and grow in biological 
systems strongly influenced by genetics and environmental inputs.\footnote{Possible 
biological implementations of the various 
sensors and sensory processes discussed in this chapter 
can be found in \cite{Churchland:1992}. The 
nature and neurophysiological implementation of visual
processes are discussed in detail in \cite{Zeki:1993}.}

@@

\subsection{Example channels}

The segment characteristics which will be 
used later in various experiments are defined below. 
Their values are all derived by straightforward computations from 
the raw image maps captured by the cameras. 
\begin{itemize} 
\item AREA: The surface area of a segment
is calculated by simply counting the number of pixels that are part 
of the segment. 
\item HPOS, VPOS: The x and y values of the 
central-position of a segment. The central position 
is calculated by taking the mid-point of the sides 
of the bounding box. 
\item HEIGHT: The height of the bounding box. 
\item WIDTH: The width of the bounding box. 
\item BB-AREA: The area of the bounding box,
calculated by multiplying height by width. 
\item GRAY: The average gray-scale value of the pixels
in a segment. 
\item R, G, B: The average R (redness), G
(greenness), and B (blueness) values in a
segment. To obtain more human-like colour channels, they are
transformed in terms of YB (yellow-blue), RG (red-green), 
BW (black-white), SAT (saturation) and BRIGHTNESS channels. 
\item EDGE-COUNT: The number of edges in
a segment, for example 3 in the case of 
a triangle. 
\item ANGLE-COUNT: The number of angles, determined on the 
basis of the junctions. 
\item RATIO: The ratio between the area of the segment 
and the area of its bounding box, which gives an idea how close 
a figure is to a rectangular shape. 
\end{itemize}

Each of these segment characteristics or combinations of
characteristics enables certain types of
categorisations. For example, the GRAY channel makes 
it possible to distinguish between 
light and dark, HEIGHT between short and tall, HPOS between 
left and right, VPOS between top and bottom, 
AREA between big and small, etc. In the next chapter, we will study 
how such categorisations may 
form on the basis of their respective channels.

@@

Here are some other examples: 
\begin{enumerate}
\item A set of sensory channels that are sensitive 
to characteristics of moving segments can easily be 
imagined. They include the speed
of movement, the direction of movement (along the 
horizontal position (HPOS) and vertical position (VPOS) 
dimensions), or the change in area (getting bigger if the object 
approaches or smaller if it moves away). This is the foundation
for categories about spatial change: moving left versus moving right,
approaching versus retracting, speeding up versus slowing down.\footnote{Experiments
using segmentation based on movement
using the dedicated hardware shown in \ref{f:plate6} were carried out by Tony Belpaeme. 
More details can be found in: \cite{Belpaeme:1998} These channels have been used in
language experiments where the agents were 
perceiving and communicating about moving 
images: \cite{Steels:1998}  
The segmentation used in this book 
based on output from the Sony EVI-D31 camera was 
implemented by Danny Van Tieghem. Angus McIntyre integrated
the interfaces to this camera within the \textsc{babel} environment. 
}

@@

\item More properties of the shape of segments can be computed.
I have already introduced the RATIO channel, which is the area of the 
segment divided by the area of the bounding box, giving an 
indication how rectangular a segment is. This can be generalised
by using an ellips around a shape so that we can calculate
convexity. We can then also calculate elongation (by calculating
the principal axes of the ellipse). 
\end{enumerate}

@@

A first frame of reference is based on
the absolute minima and maxima of the values 
on a sensory channel. This can be used to do 
{\itshape sensor-oriented scaling}. For example, 
the image map captured by the camera
contains 320 $\times$ 240 pixels, which means 
that the horizontal position (HPOS) has a value between
0 and 320 and the vertical position (VPOS) has a value 
between 0 and 240. Both values can be scaled to fall between 
0.0 and 1.0 so that they become uniform with respect to 
other sensory channels. Given a value $x$ and a 
$min$ and $max$ value, then the scaled value is 
$x' = {(x - min)}/{(max - min)}$. 
Thus the sensed value of HPOS=200 becomes
HPOS=0.62 after scaling .

In the experiments discussed later, 
sensor-scaling is always performed so that all sensory data
is between 0.0 and 1.0, allowing values on different channels 
to be compared with 
each other. In addition, context-oriented scaling is performed.
{\itshape Context-oriented scaling} uses the minima and maxima of the 
values that effectively appear in the context.
For example, if the values for the HEIGHT of segments 
observed in the scene is between 500 and 700, 
then these can be the minimum and maximum of the 
HEIGHT scale, so that 500 maps onto 0.0 and 700 to 1.0. 
Context-oriented scaling acts like a lens
magnifying the perceived values so that distinctions
stand out more clearly. This scaling could 
be made more sophisticated by introducing an additional scaling factor so that 
differences are amplified but not necessarily blown up to their extremes.

@@

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.45\textwidth]{chap3/figs/recscene.pdf}}
\caption{\label{rect1b}A scene with three objects. The
sensory values on the grayscale channel are the most salient 
and will therefore be chosen preferentially for categorisation
and verbalisation.}
\end{figure}

The saliency of a channel 
is the smallest distance (after sensor-scaling) between the perceived
values of the topic and one of the corresponding perceived values of the 
other segments in the context. 
Thus the different perceived values (after sensor-scaling)
for the segments in \figref{rect1b} are shown in 
\tabref{tab:rect1b}. The last line shows the saliency, assuming that 
the topic is segment 1.  Clearly the GRAY-channel is the most salient
one in this case, followed by the WIDTH-channel. Other 
sensory channels such as HPOS, VPOS or HEIGHT (which have almost the same 
values for object 0 and 1) are not salient at all.

\begin{table}
\begin{center}
\begin{tabular}{  l   l   l   l   l   l   l  }
\lsptoprule
{\itshape obj} & HPOS & VPOS & HEIGHT & WIDTH & GRAY & AREA \\ \midrule
0 & 0.66 & 0.95 & 0.01 & 0.71 & 0.19 & 0.27\\ 
1 & 0.69 & 0.83 & 0.07 & 0.33 & 0.97 & 0.21\\ 
2 & 0.99 & 0.87 & 0.54 & 0.72 & 0.22 & 0.57\\ 
sal & 0.03 & 0.05 & 0.07 & 0.39 & 0.75 & 0.06 \\ 
\lspbottomrule
\end{tabular}
\caption{Perceived values for each segment in \figref{rect1b}.
\label{tab:rect1b}}
\end{center}
\end{table}

Determining saliency has to be done before context scaling, because
context-scaling stretches the values to their extremes and so 
saliency information is lost. 
\tabref{tab:rect1b-area} shows for the scene shown in
\figref{rect1b} the AREA channel with its
raw data, the value after 
sensor-scaling (with minimum 1,000 and maximum 
10,000), and the value after context-scaling.

\begin{table}
\begin{center}
\begin{tabular}{  l   c   c   c  } \lsptoprule
{\itshape Object} & {\itshape Raw data} & {\itshape Sensor-scaled} &  {\itshape Context-scaled} \\ \midrule
0 & 24530 & 0.27 & 0.18 \\ 
1 & 18924 & 0.21 & 0.0\\ 
2 & 50960 & 0.57 & 1.00 \\ 
\lspbottomrule
\end{tabular}
\caption{Data for the area channel for the scene in \figref{rect1b}.
\label{tab:rect1b-area}}
\end{center}
\end{table}

\begin{figure}
\begin{center}
\includegraphics[width=0.8\columnwidth]{chap7/figs/plate-10.pdf}
\end{center}
\caption{Three examples of segmented images. The 
topic is indicated by a dashed bounding box in the 
image of the speaker. Segments which are too small 
are ignored. The topics have all been conceptualised
as being to the right and so the same word 
`gofubo' has been used to refer to them. }
\label{f:plate-10}
\end{figure}

@@

\begin{table}
\begin{center}
\begin{tabular}{  l   l   l   c  } \lsptoprule
{\itshape channel}& {\itshape obj-0} & {\itshape obj-1} & {\itshape saliency}\\ \midrule
HPOS & 0.27 & 0.16 & 0.11\\ 
VPOS & 0.20 & 0.20 & 0.0\\ 
HEIGHT & 0.15 & 0.15 & 0.0\\ 
WIDTH & 0.10 & 0.11 & 0.01\\ 
AREA & 0.10 & 0.10 & 0.0\\ 
R & 0.23 & 0.25 & 0.02\\ 
G & 0.32 & 0.34 & 0.02\\ 
B & 0.63 & 0.65 & 0.02\\ 
\lspbottomrule
\end{tabular}
\caption{Perceived values for each segment in \figref{f:plate-10}
\label{tab:t-plate10}}
\end{center}
\end{table}

The table shows clearly that HPOS is the most salient channel. 
The horizontal position also strikes us immediately 
as being the most salient when looking at  \figref{f:plate-10}. 
For many of the other channels, the differences are
almost insignificant. The use of saliency facilitates 
enormously communication and the acquisition of new categories. 
The case shown in \figref{f:plate-10} (top images) is an excellent opportunity 
for the agents to learn about left versus right. If more 
channels are salient, it is no longer so easy for the 
hearer to guess what meaning might have been used by the 
speaker and so incoherence would slip into the group's lexical
system.

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.65\textwidth]{chap3/figs/scaling.pdf}}
\caption{\label{scaling}Processing converts data on 
a sensory channel into more usable data for later categorisation processes.}
\end{figure}

@@

I hope the reader now has a much better idea of the enormous challenge
that the Talking Heads face when trying to play a language 
game about a real world scene, particularly
because they try to develop a lexicon and ontology 
as well. The images contain a multitude of ways 
to make distinctions and they differ slightly for both agents. 
It is enough fo a cloud to pass by causing the light conditions
to change slightly, and different values are immediately seen 
on the colour channels possibly leading to different segmentations.
So how can the agents ever get a repertoire of abstract categories and 
associated words when the real world shows such a perplexing 
variation? Very different scenes (for example the ones contained
in \figref{f:plate-10}) can intuitively be distinguished with the same 
categories (namely [LEFT] versus [RIGHT]). But how can such 
inductive leaps be made? Scaling and sensor transformation 
introduce some perceptual constancy, and saliency helps to 
restrict the attention to those
channels that are potentially effective in communication, but there 
is clearly an enormous gap between sensory data and language.

@@

\section{The GEOM world}

\is{GEOM world}One way to perform controlled experiments 
quickly and on a large scale consists of artificially 
generating the perceptual input to the agents. 
This has the additional advantage that we can 
simplify the situations the agents have to work in
(for example have fewer types of shapes) 
and let them start initially with a shared perception. A 
simulation world that we have used extensively 
for many simulations reported later 
is the GEOM world.

The GEOM world generates geometric shapes similar 
to the physical figures we paste on the white board. 
Possible geometric shapes are circles, triangles, 
squares, rectangles, etc. 
We can control the complexity
of the scenes that are generated  through a few parameters, for example the 
number of minimum or maximum figures, or the possible 
repertoire of shapes. We 
ignore colour and use only different grayscale 
values. To construct a scene the computer simulation 
program chooses first a random number of 
figures. Then for each figure, a shape is 
chosen, and random values for the main 
characteristics (position, height, width, grayscale)
are set.

\begin{figure}[htbp]
  \centerline{\includegraphics[width=.45\textwidth]{chap3/figs/scene9.pdf}}
\caption{\label{geom3}An example of the 
computer-generated scenes from the GEOM world.
Each shape is labeled for further reference.}
\end{figure}

@@

\begin{table}
\begin{center}
\begin{tabular}{  l   l   l   l   l   l   l   l  }
\lsptoprule
 & {\itshape HPOS} & {\itshape VPOS} & {\itshape HEIGHT} & {\itshape WIDTH} & {\itshape GRAY} & {\itshape RATIO}  & {\itshape AREA} \\ \midrule
0 & 116 & 166 & 293 & 293 & 0.777 & 1.0 & 85849 \\ 
1 & 692 & 317 & 148 & 224 & 0.449 & 1.0 & 33152 \\ 
2 & 192 & 64 & 137 & 137 & 0.408 & 1.0 & 18769 \\ 
3 & 167 & 770 & 277 & 277 & 0.201 & 1.0 & 76729 \\ 
\lspbottomrule
\end{tabular}
\end{center}
\caption{\label{tab:t-geom}Sensory data for the scene shown in \figref{geom3}.}
\end{table}

\begin{table}
\begin{center}
\begin{tabular}{  l   l   l   l   l   l   l   l  }
\lsptoprule
& {\itshape HPOS} & {\itshape VPOS} & {\itshape HEIGHT} & {\itshape WIDTH} & {\itshape GRAY} & {\itshape RATIO}  & {\itshape AREA} \\ \midrule
0 & 0.09 & 0.13 &  0.64 &  0.64 &  0.78 &  1.0 & 0.95 \\ 
1 & 0.55 & 0.25 & 0.16 & 0.41 & 0.45 & 1.0 & 0.37 \\ 
2 & 0.15 & 0.05 & 0.12 & 0.12 & 0.41 & 1.0 & 0.21 \\ 
3 & 0.13 & 0.62 & 0.59 & 0.59 & 0.20 & 1.0 & 0.85 \\ 
\lspbottomrule
\end{tabular}
\end{center}
\caption{\label{tab:t-geomscaled}Sensory data after scaling for the scene shown in \figref{geom3}.}
\end{table}

>>>>

\enlargethispage{1\baselineskip}
So we find that the world does not present itself neatly segmented, 
processed, and categorised. There is an enormous 
gap between the symbolic world of objects and categories
and the subsymbolic world of analog sensing. 
The brain somehow performs a vast amount of processing
to fill this gap, without us being in the least aware of it. 
This processing takes place in parallel. Many different
segmentation procedures and feature detectors 
operate concurrently on streams of consecutive images, 
generating hypotheses that become stronger if they are
confirmed by additional evidence, or weaker if 
they do not fit into a larger picture. We are only 
aware of the final result and therefore have no 
intuition of what is really going on, except
in rare and pathological circumstances. 
Rather than viewing perception as a straight-forward step-by-step
transformation of the raw sensory data into a segmented
picture, it is better to think of the whole process
as a boiling soup with thousands of hypotheses taking
shape, some of them floating like bubbles up to the surface. Constraints
and expectations flow down from the top so that the maximum 
amount of available information is used to construct the 
coherent segmented picture of reality we consciously experience.

@@

Sensors and actuators are the interfaces between the physical
world and the internal world.\is{sensors}\is{actuators} 
They are dedicated hard-wired components which grow in biological 
systems strongly influenced by genetics and environmental inputs.\footnote{Possible 
biological implementations of the various 
sensors and sensory processes discussed in this chapter 
can be found in \cite{Churchland:1992}. The 
nature and neurophysiological implementation of visual
processes are discussed in detail in \cite{Zeki:1993}.}

@@

\subsection{Example channels}
\enlargethispage{1\baselineskip}
The segment characteristics which will be 
used later in various experiments are defined below. 
Their values are all derived by straightforward computations from 
the raw image maps captured by the cameras. 
\begin{itemize} 
\item \textsc{area}: The surface area of a segment
is calculated by simply counting the number of pixels that are part 
of the segment. 
\item \textsc{hpos}, \textsc{vpos}: The x and y values of the 
central-position of a segment. The central position 
is calculated by taking the mid-point of the sides 
of the bounding box. 
\item \textsc{height}: The height of the bounding box. 
\item \textsc{width}: The width of the bounding box. 
\item \textsc{bb-area}: The area of the bounding box,
calculated by multiplying height by width. 
\item \textsc{gray}: The average gray-scale value of the pixels
in a segment. 
\item R, G, B: The average R (redness), G
(greenness), and B (blueness) values in a
segment. To obtain more human-like colour channels, they are
transformed in terms of YB (yellow-blue), RG (red-green), 
BW (black-white), SAT (saturation) and \textsc{brightness} channels. 
\item \textsc{edge-count}: The number of edges in
a segment, for example 3 in the case of 
a triangle. 
\item \textsc{angle-count}: The number of angles, determined on the 
basis of the junctions. 
\item \textsc{ratio}: The ratio between the area of the segment 
and the area of its bounding box, which gives an idea how close 
a figure is to a rectangular shape. 
\end{itemize}

Each of these segment characteristics or combinations of
characteristics enables certain types of
categorisations. For example, the \textsc{gray} channel makes 
it possible to distinguish between 
light and dark, \textsc{height} between short and tall, \textsc{hpos} between 
left and right, \textsc{vpos} between top and bottom, 
\textsc{area} between big and small, etc. In the next chapter, we will study 
how such categorisations may 
form on the basis of their respective channels.

@@

\clearpage
Here are some other examples: 
\begin{enumerate}
\item A set of sensory channels that are sensitive 
to characteristics of moving segments can easily be 
imagined. They include the speed
of movement, the direction of movement (along the 
horizontal position (\textsc{hpos}) and vertical position (\textsc{vpos}) 
dimensions), or the change in area (getting bigger if the object 
approaches or smaller if it moves away). This is the foundation
for categories about spatial change: moving left versus moving right,
approaching versus retracting, speeding up versus slowing down.\footnote{Experiments
using segmentation based on movement
using the dedicated hardware shown in \ref{f:plate6} were carried out by Tony Belpaeme. 
More details can be found in: \cite{Belpaeme:1998} These channels have been used in
language experiments where the agents were 
perceiving and communicating about moving 
images: \cite{Steels:1998}  
The segmentation used in this book 
based on output from the Sony EVI-D31 camera was 
implemented by Danny Van Tieghem. Angus McIntyre integrated
the interfaces to this camera within the \textsc{babel} environment. 
}

@@

\item More properties of the shape of segments can be computed.
I have already introduced the \textsc{ratio} channel, which is the area of the 
segment divided by the area of the bounding box, giving an 
indication how rectangular a segment is. This can be generalised
by using an ellips around a shape so that we can calculate
convexity. We can then also calculate elongation (by calculating
the principal axes of the ellipse). 
\end{enumerate}

@@

A first frame of reference is based on
the absolute minima and maxima of the values 
on a sensory channel. This can be used to do 
{\itshape sensor-oriented scaling}. For example, 
the image map captured by the camera
contains 320 × 240 pixels, which means 
that the horizontal position (\textsc{hpos}) has a value between
0 and 320 and the vertical position (\textsc{vpos}) has a value 
between 0 and 240. Both values can be scaled to fall between 
0.0 and 1.0 so that they become uniform with respect to 
other sensory channels. Given a value $x$ and a 
$min$ and $max$ value, then the scaled value is 
$x' = {(x - min)}/{(max - min)}$. 
Thus the sensed value of \textsc{hpos}=200 becomes
\textsc{hpos}=0.62 after scaling .

In the experiments discussed later, 
sensor-scaling is always performed so that all sensory data
is between 0.0 and 1.0, allowing values on different channels 
to be compared with 
each other. In addition, context-oriented scaling is performed.
{\itshape Context-oriented scaling} uses the minima and maxima of the 
values that effectively appear in the context.
For example, if the values for the \textsc{height} of segments 
observed in the scene is between 500 and 700, 
then these can be the minimum and maximum of the 
\textsc{height} scale, so that 500 maps onto 0.0 and 700 to 1.0. 
Context-oriented scaling acts like a lens
magnifying the perceived values so that distinctions
stand out more clearly. This scaling could 
be made more sophisticated by introducing an additional scaling factor so that 
differences are amplified but not necessarily blown up to their extremes.

@@

\begin{figure}[h]
  \centerline{\includegraphics[width=.45\textwidth]{chap3/figs/recscene.pdf}}
\caption{\label{rect1b}A scene with three objects. The
sensory values on the grayscale channel are the most salient 
and will therefore be chosen preferentially for categorisation
and verbalisation.}
\end{figure}

The saliency of a channel 
is the smallest distance (after sensor-scaling) between the perceived
values of the topic and one of the corresponding perceived values of the 
other segments in the context. 
Thus the different perceived values (after sensor-scaling)
for the segments in \figref{rect1b} are shown in 
\tabref{tab:rect1b}. The last line shows the saliency, assuming that 
the topic is segment 1.  Clearly the \textsc{gray}-channel is the most salient
one in this case, followed by the \textsc{width}-channel. Other 
sensory channels such as \textsc{hpos}, \textsc{vpos} or \textsc{height} (which have almost the same 
values for object 0 and 1) are not salient at all.

Determining saliency has to be done before context scaling, because
context-scaling stretches the values to their extremes and so 
saliency information is lost. 
\tabref{tab:rect1b-area} shows for the scene shown in
\figref{rect1b} the \textsc{area} channel with its
raw data, the value after 
sensor-scaling (with minimum 1,000 and maximum 
10,000), and the value after context-scaling.

\begin{table}[b]
\begin{center}
\begin{tabular}{  l   l   l   l   l   l   l  }
\lsptoprule
{\itshape obj} & \textsc{hpos} & \textsc{vpos} & \textsc{height} & \textsc{width} & \textsc{gray} & \textsc{area} \\ \midrule
0 & 0.66 & 0.95 & 0.01 & 0.71 & 0.19 & 0.27\\ 
1 & 0.69 & 0.83 & 0.07 & 0.33 & 0.97 & 0.21\\ 
2 & 0.99 & 0.87 & 0.54 & 0.72 & 0.22 & 0.57\\ 
sal & 0.03 & 0.05 & 0.07 & 0.39 & 0.75 & 0.06 \\ 
\lspbottomrule
\end{tabular}
\caption{Perceived values for each segment in \figref{rect1b}.
\label{tab:rect1b}}
\end{center}
\end{table}

\begin{table}[t]
\begin{center}
\begin{tabular}{  l   c   c   c  } \lsptoprule
{\itshape Object} & {\itshape Raw data} & {\itshape Sensor-scaled} &  {\itshape Context-scaled} \\ \midrule
0 & 24530 & 0.27 & 0.18 \\ 
1 & 18924 & 0.21 & 0.0\\ 
2 & 50960 & 0.57 & 1.00 \\ 
\lspbottomrule
\end{tabular}
\caption{Data for the area channel for the scene in \figref{rect1b}.
\label{tab:rect1b-area}}
\end{center}
\end{table}

\begin{figure}[hbtp]
\begin{center}
\includegraphics[width=0.8\columnwidth]{chap7/figs/plate-10.pdf}
\end{center}
\caption{Three examples of segmented images. The 
topic is indicated by a dashed bounding box in the 
image of the speaker. Segments which are too small 
are ignored. The topics have all been conceptualised
as being to the right and so the same word 
`gofubo' has been used to refer to them. }
\label{f:plate-10}
\end{figure}

@@

\begin{table}[t]
\begin{center}
\begin{tabular}{  l   l   l   c  } \lsptoprule
{\itshape channel}& {\itshape obj-0} & {\itshape obj-1} & {\itshape saliency}\\ \midrule
\textsc{hpos} & 0.27 & 0.16 & 0.11\\ 
\textsc{vpos} & 0.20 & 0.20 & 0.0\\ 
\textsc{height} & 0.15 & 0.15 & 0.0\\ 
\textsc{width} & 0.10 & 0.11 & 0.01\\ 
\textsc{area} & 0.10 & 0.10 & 0.0\\ 
R & 0.23 & 0.25 & 0.02\\ 
G & 0.32 & 0.34 & 0.02\\ 
B & 0.63 & 0.65 & 0.02\\ 
\lspbottomrule
\end{tabular}
\caption{Perceived values for each segment in \figref{f:plate-10}
\label{tab:t-plate10}}
\end{center}
\end{table}

The table shows clearly that \textsc{hpos} is the most salient channel. 
The horizontal position also strikes us immediately 
as being the most salient when looking at  \figref{f:plate-10}. 
For many of the other channels, the differences are
almost insignificant. The use of saliency facilitates 
enormously communication and the acquisition of new categories. 
The case shown in \figref{f:plate-10} (top images) is an excellent opportunity 
for the agents to learn about left versus right. If more 
channels are salient, it is no longer so easy for the 
hearer to guess what meaning might have been used by the 
speaker and so incoherence would slip into the group's lexical
system.

\begin{figure}[t]
  \centerline{\includegraphics[width=.65\textwidth]{chap3/figs/scaling.pdf}}
\caption{\label{scaling}Processing converts data on 
a sensory channel into more usable data for later categorisation processes.}
\end{figure}

@@

I hope the reader now has a much better idea of the enormous challenge
that the Talking Heads face when trying to play a language 
game about a real world scene, particularly
because they try to develop a lexicon and ontology 
as well. The images contain a multitude of ways 
to make distinctions and they differ slightly for both agents. 
It is enough fo a cloud to pass by causing the light conditions
to change slightly, and different values are immediately seen 
on the colour channels possibly leading to different segmentations.
So how can the agents ever get a repertoire of abstract categories and 
associated words when the real world shows such a perplexing 
variation? Very different scenes (for example the ones contained
in \figref{f:plate-10}) can intuitively be distinguished with the same 
categories (namely [\textsc{left}] versus [\textsc{right}]). But how can such 
inductive leaps be made? Scaling and sensor transformation 
introduce some perceptual constancy, and saliency helps to 
restrict the attention to those
channels that are potentially effective in communication, but there 
is clearly an enormous gap between sensory data and language.

@@

\section{The \textsc{geom} world}

\is{\textsc{geom} world}One way to perform controlled experiments 
quickly and on a large scale consists of artificially 
generating the perceptual input to the agents. 
This has the additional advantage that we can 
simplify the situations the agents have to work in
(for example have fewer types of shapes) 
and let them start initially with a shared perception. A 
simulation world that we have used extensively 
for many simulations reported later 
is the \textsc{geom} world.

The \textsc{geom} world generates geometric shapes similar 
to the physical figures we paste on the white board. 
Possible geometric shapes are circles, triangles, 
squares, rectangles, etc. 
We can control the complexity
of the scenes that are generated  through a few parameters, for example the 
number of minimum or maximum figures, or the possible 
repertoire of shapes. We 
ignore colour and use only different grayscale 
values. To construct a scene the computer simulation 
program chooses first a random number of 
figures. Then for each figure, a shape is 
chosen, and random values for the main 
characteristics (position, height, width, grayscale)
are set.

\begin{figure}[t]
  \centerline{\includegraphics[width=.45\textwidth]{chap3/figs/scene9.pdf}}
\caption{\label{geom3}An example of the 
computer-generated scenes from the \textsc{geom} world.
Each shape is labeled for further reference.}
\end{figure}

@@

\begin{table}[b]
\begin{center}
\begin{tabular}{  l   l   l   l   l   l   l   l  }
\lsptoprule
 & {\itshape \textsc{hpos}} & {\itshape \textsc{vpos}} & {\itshape \textsc{height}} & {\itshape \textsc{width}} & {\itshape \textsc{gray}} & {\itshape \textsc{ratio}}  & {\itshape \textsc{area}} \\ \midrule
0 & 116 & 166 & 293 & 293 & 0.777 & 1.0 & 85849 \\ 
1 & 692 & 317 & 148 & 224 & 0.449 & 1.0 & 33152 \\ 
2 & 192 & 64 & 137 & 137 & 0.408 & 1.0 & 18769 \\ 
3 & 167 & 770 & 277 & 277 & 0.201 & 1.0 & 76729 \\ 
\lspbottomrule
\end{tabular}
\end{center}
\caption{\label{tab:t-geom}Sensory data for the scene shown in \figref{geom3}.}
\end{table}

\begin{table}
\begin{center}
\begin{tabular}{  l   l   l   l   l   l   l   l  }
\lsptoprule
& {\itshape \textsc{hpos}} & {\itshape \textsc{vpos}} & {\itshape \textsc{height}} & {\itshape \textsc{width}} & {\itshape \textsc{gray}} & {\itshape \textsc{ratio}}  & {\itshape \textsc{area}} \\ \midrule
0 & 0.09 & 0.13 &  0.64 &  0.64 &  0.78 &  1.0 & 0.95 \\ 
1 & 0.55 & 0.25 & 0.16 & 0.41 & 0.45 & 1.0 & 0.37 \\ 
2 & 0.15 & 0.05 & 0.12 & 0.12 & 0.41 & 1.0 & 0.21 \\ 
3 & 0.13 & 0.62 & 0.59 & 0.59 & 0.20 & 1.0 & 0.85 \\ 
\lspbottomrule
\end{tabular}
\end{center}
\caption{\label{tab:t-geomscaled}Sensory data after scaling for the scene shown in \figref{geom3}.}
\end{table}
